---
title: "Module 4 Pair Quiz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Ayuson, Marc Edison

See, Sean Jacob

7/29/21

## Questions {.tabset .tabfade}

### Question 1

**1. An article in the Journal of Sound and Vibration ["Measurement of Noise-Evoked Blood Pressure by Means of Averaging Method: Relation between Blood Pressure Rise and PSL" (1991, Vol. 151(3), pp. 383-394)] described a study investigating the relationship between noise exposure and hypertension. The following data are representative of those reported in the article.**

---

#### A. Draw a scatter diagram of y (blood pressure rise in millimeters of mercury) versus x (sound pressure level in decibels). Does a simple linear regression model seem reasonable in this situation?

```{r}
y <- c(1, 0, 1, 2, 5, 1, 4, 6, 2, 3, 5, 4, 6, 8, 4, 5, 7, 9, 7, 6)

x <- c(60,	63,	65,	70,	70,	70,	80,	90,	80,	80,	85,	89,	90,	90,	90,	90,	94,	100,	100,	100)

plot(x, y, pch=16, cex=1, main = "Sound Pressure Level(dB) vs Blood Pressure Rise(mmHg)" ,xlab = "Sound Pressure Level in Decibels (dB)", ylab= "Blood Pressure Rise in Millimeters of Mercury (mmHg)")

lm(y ~ x)

abline(lm(y ~ x), col = "blue")
```

---

#### B. Fit the simple linear regression model using least squares. Find an estimate of $\sigma^2$.

```{r}

fit <- lm(y ~ x)

summary(fit)

(summary(fit)$sigma)**2 

```

---

#### C. Find the predicted mean rise in blood pressure level associated with a sound pressure level of 85 decibels.

```{r}

Sxy = sum((x - mean(x)) * (y - mean(y)))
Sxx = sum((x - mean(x)) ^ 2)
Syy = sum((y - mean(y)) ^ 2)
c(Sxy, Sxx, Syy)

beta_1_hat = Sxy / Sxx
beta_0_hat = mean(y) - beta_1_hat * mean(x)
c(beta_0_hat, beta_1_hat)

```